---
title: "Assignment2_DataAnalysis"
format: html
---

## Project Summary & Instructions

EPI 203 Assignment 2. This project involves analyzing a synthetic dataset with any prediction model.

### Instructions

Create a short R analysis of the cohort simulated data in a repository on your GitHub account. Fork the repository and then edit to add your analysis in a new branch. The analysis should include at least one table describing the variables, a prediction algorithm (e.g., glm function, caret package, etc.), at least one figure, and a brief summary of your findings in the README.

## Data Cleaning

Data set: synthetic data

The csv file for `cohort` in the `raw-data` folder includes 5,000 observations with variables `smoke`, `female`, `age`, `cardiac`, and `cost`.

```{r}
library(tidyverse)
library(here)

# load dataset
synthetic_df <- read_csv("/Users/tracychidyausiku/Coursework/epi/epi203/Assignment2/raw-data/cohort.csv")

head(synthetic_df)

# check for any missingness
sum(is.na(synthetic_df))

# visualize distribution of continuous variables
hist(synthetic_df$age)
hist(synthetic_df$cost)
```

## Table 1

Purpose: Create a table with numeric description of variables

Binary Variables: female, cardiac, smoke

Continuous Variables: age, cost

```{r}
library(table1)

# Make binary variables factors
synthetic_df$female <- factor(synthetic_df$female, levels=c(0,1), labels=c("Male", "Female"))
synthetic_df$smoke <- factor(synthetic_df$smoke, levels=c(0,1), labels=c("No Smoke", "Smoke"))
synthetic_df$cardiac <- factor(synthetic_df$cardiac, levels=c(0,1), labels=c("No Cardiac", "Cardiac"))

# Create table labels
label(synthetic_df$female) <- "Sex"
label(synthetic_df$smoke) <- "Smoke"
label(synthetic_df$cost) <- "Cost"
label(synthetic_df$age) <- "Age"

# create table 1
table1(~ cost + age + female + smoke | cardiac, data = synthetic_df)
```

## 
