---
title: "Assignment2_DataAnalysis"
format: gfm
---

## Project Summary & Instructions

EPI 203 Assignment 2. This project involves analyzing a synthetic dataset with any prediction model.

### Instructions

Create a short R analysis of the cohort simulated data in a repository on your GitHub account. Fork the repository and then edit to add your analysis in a new branch. The analysis should include at least one table describing the variables, a prediction algorithm (e.g., glm function, caret package, etc.), at least one figure, and a brief summary of your findings in the README.

## Data Cleaning

Data set: synthetic data

The csv file for `cohort` in the `raw-data` folder includes 5,000 observations with variables `smoke`, `female`, `age`, `cardiac`, and `cost`.

```{r}
library(tidyverse)
library(here)

# load dataset
synthetic_df <- read_csv("/Users/tracychidyausiku/Coursework/epi/epi203/Assignment2/raw-data/cohort.csv")

head(synthetic_df)

# check for any missingness
sum(is.na(synthetic_df))

# visualize distribution of continuous variables
hist(synthetic_df$age)
hist(synthetic_df$cost)
```

## Table 1

Purpose: Create a table with numeric description of variables

Binary Variables: female, cardiac, smoke

Continuous Variables: age, cost

```{r}
library(table1)
library(kableExtra)

# Make binary variables factors
synthetic_df$female <- factor(synthetic_df$female, levels=c(0,1), labels=c("Male", "Female"))
synthetic_df$smoke <- factor(synthetic_df$smoke, levels=c(0,1), labels=c("No Smoke", "Smoke"))
synthetic_df$cardiac <- factor(synthetic_df$cardiac, levels=c(0,1), labels=c("No Cardiac", "Cardiac"))

# Create table labels
label(synthetic_df$female) <- "Sex"
label(synthetic_df$smoke) <- "Smoke"
label(synthetic_df$cost) <- "Cost"
label(synthetic_df$age) <- "Age"

# create table 1
tableOne <- table1(~ cost + age + female + smoke | cardiac, data = synthetic_df)

# Save as PNG
save_kable(tableOne, file = "table1.png")
```

## Data Analysis

Question: Does cost influence the risk for a cardiac event?

Exposure: cost

Outcome: cardiac (occurs in \~5% of the population)

Covariates: female, smoke, age

Model: Logistic Regression

```{r}
# model: logistic regression
logistic_reg <- glm(cardiac ~ cost + female + smoke + age, data = synthetic_df, family = "binomial")

summary(logistic_reg)

# OR (CI) table
results <- exp(cbind(OR = coef(logistic_reg), confint(logistic_reg)))

# Save ORs (CI) as a dataframe
logistc_reg_OR_CI <- as.data.frame(results)
```

##  

## Figures

```{r}
library(ggplot2)

table(synthetic_df$female, synthetic_df$smoke)

# smoker by female/male
ggplot(synthetic_df, aes(x = smoke, fill = female)) +
  geom_bar(alpha = 0.5, position = "identity", bins = 30) +
  scale_fill_manual(values = c("Female" = "red", "Male" = "blue")) +
  labs(title = "Smoke Distribution by Gender",
       x = "Smoke",
       y = "Frequency")

# smoker by age
ggplot(synthetic_df, aes(x=age, fill=smoke)) +
  geom_histogram(binwidth=5, position="identity", alpha=0.5) +
  labs(title="Age Distribution by Smoking Status",
       x="Age",
       y="Frequency") +
  theme_minimal()
```

### Forest Plot

```{r}
library(finalfit)

explanatory = c("age", "female", "smoke", "cost")
dependent = 'cardiac'
synthetic_df %>% or_plot(dependent, explanatory, 
                         dependent_label = "Forest Plot, Cardiac")

```
